---
title: "HW 03: Multiple Linear Regression"
author: "Hannah Wang"
date: "2019 October 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE,
                      message=FALSE)
```

```{r load-packages}
library(tidyverse)
library(broom)
library(knitr) 
```

```{r load-data}
# for Question 8
houses <- read_csv("data/KingCountyHouses.csv")
```

### Question 1

speed-hat = 52.387 + 0.020 * year - 0.0032 * starters - 1.070	* good -2.1834 * slow + 0.012 * (year x good) + 0.012 * (year x slow)

### Question 2

For slow  track conditions: 

speed-hat = 52.387 + 0.020 * year - 0.0032 * starters - 1.070	* 0 - 2.1834 * 1 + 0.012 * (year x 0) + 0.012 * (year x 1)

= 50.2036 + 0.032 * year - 0.0032 * starters

### Question 3

The intercept (52.387 feet per second) describes the average speed for a winner in the year 1896, with a fast track condition, and racing with 0 horses. The intercept is not meaningful because it does not make sense for the winner to race against 0 horses.

### Question 4

The p-value is 0.013. We use the p-value in the conditiongood variable row because that is the p-value we get from comparing good track conditions to the baseline condition of fast track. Since year = 0, we don't have to worry about the interaction variables.

### Question 5

All other variables held constant, the winner's speed is predicted, on average, to be 2.183 feet per second lower on a slow track than on a fast track.

### Question 6

The p-value is 0.113. We use the p-value in the year:conditiongood interaction variable row.

### Question 7 


### Question 8

```{r new-variables}
houses <- houses %>%
  filter(bedrooms <= 5 ) %>%
  mutate(floorsCat = as.factor(floors), 
         sqftCent = sqft - mean(sqft), 
         bedroomsCent = bedrooms - mean(bedrooms),
         bathroomsCent = bathrooms-mean(bathrooms),
         logprice = log(price))
```




### Overall (do not delete!)


